---
title: "Rep_Research_Peer_Assessment_1"
author: "Vince Glick"
date: "Sunday, February 15, 2015"
output: html_document
---

```{r}
echo = TRUE
options(scipen = 1)
```
  
  
#Loading and preprocessing the data

```{r}
require(ggplot2)
```

##1. Load the data (i.e. read.csv() )

```{r}
activity <- read.csv("activity.csv", colClasses = c("integer", "Date", "factor"))
```

##2. Process/transform the data (if necessary) into a format suitable for your analysis

```{r}
activclean <- na.omit(activity)
activity$month <- as.numeric(format(activity$date, "%m"))
```

#What is mean total number of steps taken per day?

##1. Make a histogram of the total number of steps taken each day

```{r}
gbarplot <- ggplot(data=activclean , aes(x=date, y=steps))+ geom_bar(stat = "identity")
gbarplot
```

##2. Calculate and report the mean and median total number of steps taken per day

```{r}
aggregate_steps <- aggregate(activclean$steps, list(date = activclean$date), FUN = "sum")$x

meanAvgSteps <- mean(aggregate_steps)
meanAvgSteps

medianAvgSteps <- median(aggregate_steps)
medianAvgSteps
```

#What is the average daily activity pattern?

##1. Make a time series plot (i.e. type = "l" ) of the 5minute interval (xaxis) 
##and the average number of steps taken, averaged across all days (yaxis)

```{r}
avgsteps <- aggregate(activclean$steps, list(interval = as.numeric(as.character(activclean$interval))), 
                      FUN = "mean")
names(avgsteps)[2] <- "meanAvgSteps"

ggplot(avgsteps, aes(interval, meanAvgSteps)) + 
  geom_line(color = "#d41e1e", size = 0.5) + 
  labs(title = "Time Series Plot", x = "5 Min. intervals", y = "AVG # Steps")

head(activclean)
```

##2. Which 5-minute interval, on average across all the days in the dataset, 
##contains the maximum number of steps?

```{r}
avgsteps[avgsteps$meanAvgSteps==max(avgsteps$meanAvgSteps),]
```

#Imputing missing values

##1. Calculate and report the total number of missing values in the dataset 
##(i.e. the total number of rows with NA s)

```{r}
sum(is.na(activity[1]))
```

##2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to
##be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5minute
##interval, etc.

##3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
fillActivity <- activity 
for (i in 1:nrow(fillActivity)) {
  if (is.na(fillActivity$steps[i])) {
    fillActivity$steps[i] <- avgsteps[which(fillActivity$interval[i] == avgsteps$interval), ]$meanAvgSteps
  }
}

head(fillActivity)
sum(is.na(fillActivity[1]))
```

##4. Make a histogram of the total number of steps taken each day and Calculate and report the mean
##and median total number of steps taken per day. Do these values differ from the estimates from the
##first part of the assignment?

```{r}
gbarplot2 <- ggplot(data=fillActivity, aes(x=date, y=steps))+ geom_bar(stat = "identity")
gbarplot2


avgsteps2 <- aggregate(fillActivity$steps, list(interval = as.numeric(as.character(fillActivity$interval))), 
                       FUN = "mean")$x
meanAvgSteps2 <- mean(avgsteps2)
meanAvgSteps2

medianAvgSteps2 <- median(avgsteps2)
medianAvgSteps2
```

##What is the impact of imputing missing data on the estimates of the total
##daily number of steps?

```{r}
aggregate_steps2 <- aggregate(fillActivity$steps, list(date = fillActivity$date), FUN = "sum")$x


t.test(aggregate_steps, aggregate_steps2)
wilcox.test(aggregate_steps, aggregate_steps2)
var.test(aggregate_steps, aggregate_steps2)
```


#Are there differences in activity patterns between weekdays and weekends?#

##1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating
##whether a given date is a weekday or weekend day.

```{r}
head(fillActivity)
fillActivity$day <- factor(format(fillActivity$date, "%A"))
levels(fillActivity$day)
levels(fillActivity$day) <- 
  list(weekday = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"), 
       weekend = c("Saturday", "Sunday"))
head(fillActivity)
```

##2. Make a panel plot containing a time series plot (i.e. type = "l" ) of the 5minute
##interval (xaxis) and the average number of steps taken, averaged across 
##all weekday days or weekend days (yaxis). See the README file in the GitHub repository 
##to see an example of what this plot should look like using simulated data.

```{r}
avgsteps <- aggregate(fillActivity$steps, 
                      list(interval = as.numeric(as.character(fillActivity$interval)), 
                           day = fillActivity$day), 
                      FUN = "mean")
names(avgsteps)[3] <- "meanSteps"
library(lattice)
xyplot(avgsteps$meanSteps ~ avgsteps$interval | avgsteps$day, 
       layout = c(1, 2), type = "l", 
       xlab = "INTERVAL", ylab = "# OF STEPS")
```

